\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-15 Manuscript/manuscript.tex
Started work on discussion

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. National vs International compares data from the same athletes
  who competed at 2022 national-level championships and the 2022 World Track and
  Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic & Athletes & Number observations  \\ 
   \midrule
   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\
   2022 National vs International Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
   2022 National vs International Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
   2022 vs 2023 Men & $<0.001$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
   2022 vs 2023 Female & $<0.001$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}
\of{Table currently doesnt fit on the page but can be fixed by abbreviating some
of the names in the comparison column.  I used <0.001 becuase that is what we
had currently been using for any values from the permutation test where the value
was < 2.2e-16 in R.  Professor Yan will know best, but are the sample sizes now
sufficiently large enough that the permutation test is not needed? If I recall
correctlty that was a correction for small sample sizes.}

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation and asymptotic rank-based tests for the three comparisons, 
as discussed previously in Section~\ref{sec:rank}. The national versus international
comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023comparisons.%, with the
%national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
The highly significant permutation test result provides substantial evidence 
that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior. For the 2019 versus 2022
same athlete comparison with permutation test, the p-value was 
$0.028$, which is significant at level $0.05$.  The result of
comparing 2022 and 2023 was very similar (but not exactly the same) to the above
result, as the p-value from the permutation test was $0.011$.
Taken together, these three
tests help show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed at other
competitions.

@@

We did not consider data from the women's 100 meter hurdle in the analysis.
In addition to the concerns over reaction time differences in men and
women, exploratory analysis showed that 2022 women were at best only
slightly faster relative to other championships going back to 2003.
Likewise, the rank-based comparisons resulted in p-values
that were not drastically significant after adjusting for multiple tests.
The lack of evidence from the women's data makes it 
harder to conclude that the sensors used at the 2022 championships were faulty, 
although it is unknown if the sensors for the men's 110 meter hurdle and women's
100 meter hurdle were the same. On the other hand, this may be related to 
another debated question regarding whether the reaction time barrier should be 
the same for men and women.

>>>>

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. National vs International compares data from the same athletes
  who competed at 2022 national-level championships and the 2022 World Track and
  Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic & Athletes & Number observations  \\ 
   \midrule
   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\
   2022 Nat. vs Inter. Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
   2022 Nat. vs Inter. Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
   2022 vs 2023 Men & $<0.001$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
   2022 vs 2023 Female & $<0.001$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}
\of{Table might still be too wide.  I used <0.001 becuase that is what we
had currently been using for any values from the permutation test where the value
was < 2.2e-16 in R.  Professor Yan will know best, but are the sample sizes now
sufficiently large enough that the permutation test is not needed? If I recall
correctlty that was a correction for small sample sizes.}

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation and asymptotic rank-based tests for the six comparisons, 
as discussed previously in Section~\ref{sec:rank}. The national versus international
comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023
comparisons. The highly significant permutation test results
provides substantial evidence that both men and women who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior. When looked at in context
with other tests, these results show that athletes competing at the 2022 World
Track and Field Championships had faster reaction times than when they had
competed at other competitions.

@@

We set out to answer two questions: were athletes who competed at the 2022 World
Track and Field Championships faster than at other races and is the 0.1 second
reaction time barrier fair.  In regards to the first question, we do believe
that athletes were faster than in comparable meets.  Ideally we would like to
be able to draw data from a central database containing every World Athletics
certified meet which we could use to reference athlete's reaction times.
Unfortunately that is not available at the moment and nearly all of the results
listed on the World Athletics website do not contain reaction times.  The results
of the study could have been stronger if we had been able to look at every race
that athletes like Devon Allen ran in 2022 and analyzed his results at the World
Championship against many other data points.

In addition to the GLMM results results presented above women, the supplementary
file contains results about the women's GLMM results.  We fit a generalized
gamma model to womens 100 meter dash and 100 meter hurdle data from 1999 to 2023
to determine an appropriate reaction time barrier.  We chose to separate these
two studies due to publications suggesting reaction time differences in men and
women.

\of{I think there are some good ideas in these 2 paragraphs but need help
integrating them. Prof Schifano can you help with this?}
